Coherent optical association of a single molecule

### 2020-10-08 paper.tex
Update text for figure 4

\begin{figure*}
  \includegraphics[height=5cm]{fig1.pdf}
  \caption{Optical creation of single molecule from single atoms in tweezer.
    (A) Schematics of the optical transition from an atom pair to a weakly bound molecule.
    The initial state is the relative motional ground state between the two atoms
    and the final state is the first molecular bound state.
    The transition is driven by a pair of laser frequencies matching the binding energy
    of the molecule.
    The lasers are detuned from an excited molecular state in the $c^3\Sigma$ potential
    by $\Delta$ in order to suppress the scattering during the transfer.
    (B) Comparison between using a weakly bound and a deeply bound excited state
    as intermediate state for the Raman transition.
    The deeply bound excited state (upper half $v=0$) has a smaller Raman Rabi frequency ($\Omega_{Raman}$)
    compared to the weakly bound excited state (lower half $v=63$) at a given detuning.
    However, the scattering rate ($\Gamma_{scatter}$) is also much lower,
    which results in a larger Raman Rabi frequency to scattering rate ratio.
    (C) Enhancement of short range wavefunction.
    The large scattering length for the $Na(2,2),Cs(3,3)$ state creates an interaction shift
    comparable to the axial trapping frequency.
    This causes a significant change in the relative wavefunction especially at short
    intranuclear distance ($R$).
    Compared to other spin states with weaker interaction,
    the wavefunction at short distance ($R<100\mathrm{\AA}$) is significantly enhanced.
    \label{f-theory}
  }
\end{figure*}

@@

% The optical transfer scheme we use is shown in figure \todo{}. To drive the atoms from
% We use an optical Raman transition to drive the system from the atomic initial state to a ground electronic molecular state. In order to reduce the size mismatch between the atomic and molecular states, we selected the first bound state for the 3322 spin state \todo{asymtopt to the 3322 threshold?} as our final state. \todo{move selection of intermediate state to intro?}. For similar reason, the natural choice for the intermediate excited molecular state is one with highly excited motional level.However, from our calculation (and experiment?), the smaller level space for high vibrational state and the smaller detuning from the atomic threshold increases the scattering rate of the molecular state which causes a reduced Raman Rabi frequency to decoherence rate ratio and a lower transfer efficiency. Therefore, in our experiment, we selected the v=0 state as the intermediate state for our Raman transition.

@@

To find the optimal excited intermediate state, we perform a  calculation of the Raman Rabi frequency and scattering rate at different detunings from the atomic threshold taking into account of all states of the $c^3\Sigma$ excited molecular state potential (full result in SM)~\cite{Liu2017} \textcolor{red}{KW: I have only done the calculations with c3Sigma only, cause the b3Pi is highly spin orbit coupled, and not necessarily trustworthy where we are. I can also include a citation to the paper with that potential?}. As shown in Fig.~\ref{f-theory}B, the ratio of the Raman Rabi rate to scattering rate can be made larger for more deeply bound states compared to weakly bound states at a cost of a smaller Raman Rabi frequency.  %Since for a typical molecular potential the spacing between states
% decreases with increasing vibrational levels,
% a weakly bound excited molecular state
% This further increases the scattering rate for a weakly bond excited state.
% Based on these considerations, % a weakly bound excited molecular state may not be the best choice for the intermediate state in the Raman transfer. Thus,
% we calculated both the Raman Rabi frequency and the scattering rate at different detunings (full result is shown in SM). As seen in Fig.~\ref{f-theory}B,
% we found a distinct optimal regime from what has been explored experimentally previously.
% Namely, despite a smaller Raman Rabi frequency at deeply bound intermediate states, the Rabi frequency to scattering rate ratio is much higher as compared to a weakly bound intermediate state.
As a result, we choose a deeply bound molecular excited state ($v'=0$ of $c^3\Sigma^+$) as an intermediate state to drive the Raman transition. %(mention the trend, also v=12 will work as similarly as v=0)

@@

The ratio of molecule scattering rate to the Rabi frequency is larger than the theory prediction. In order to understand the origin of this discrepancy, we measured the dependency of the Raman resonance (what does dependence of the Raman resonance mean? Do you just mean we measured molecule formation as a function of all these parameters?) as a function of the tweezer power and single photon frequency. The important results from the fits are the 2 photon resonance frequency (light shift), Raman Rabi frequency, atomic lifetime and molecular lifetime. Each of these quantities provide us with information about a different combination of physical processes.

First we look at the change in resonance frequency. As a function of the tweezer power, we observe a linear dependency on the resonance frequency caused by the differential light shift between the atomic and molecular state (Fig.~\ref{f-det}A). When we vary the tweezer frequency around the $v=0$ excited state, we can further observe a $1/\delta$ component and a background component that is constant for different detunings in the experimentally explored region. The background is caused by coupling to other excited states that are further away in energy. %which is not very important in this case.
% The $1/\delta$ component, however, is due to the coupling to the $v=0$ excited state and from equation (?) we know that the light shift is mainly due to the coupling between the excited and ground molecular states.
The $1/\delta$ component, however, is predominantly due to the coupling between the molecular state and the $ v =0 $ excited state. From this measurement, we can calculate a matrix element between the molecular state and the excited state, $ \Omega_m$, of $....$ which is similar to what we calculated from theory (ref/sm theory). (make sure theory part mentions $\Omega_{down}\gg\Omega_{up}$, also $\Omega$ vs $\Omega'$ for the single leg vs cross coupling number).
% what's the take away of this paragraph with respect to the mystery that is setup in the previous paragraph of having a larger scattering rate to rabi frequency ratio than expected?

(talk about blue side?)

Next, we discuss the dependencies of the Raman Rabi frequency. The Raman Rabi frequency shows a non-linear dependency on the tweezer power due to the change in the atomic wavefunction caused by the change in confinement (Fig.~\ref{f-det}?). The atomic matrix element, $\Omega_a$, is proportional to the the short range atomic wavefunction amplitude, which scales as $ P^{0.375} $ for weakly interacting particles. However, due to the strong interaction between the two atoms, this approximation breaks down. Instead, theory calculation shows that the scaling is very well approximated by $ P^{0.29} $ within the range of confinement in our experiment. Combined with the intensity factor, the Raman Rabi frequency should scale with $P^{1.29}$, which agrees with our experimental result. Similar to the light shift, there is also a background component and a $v=0$ component in the Raman Rabi frequency. The $v=0$ component predicts a up leg Rabi frequency of $...$ which is consistent with theory. Unfortunately, the background Raman Rabi frequency cancels the Rabi frequency for red detuning Raman transition which is one of the factors that decreases our transfer efficiency. (need blue side?)
% I think we should include some takeaway from this paragraph as well, otherwise it may not be clear what the point is. Maybe something like, "Based on these analyses, the Rabi frequencies agree with theory. Thus, the high scattering rate to raman rabi frequency ratio is caused by larger than expected scattering, which we discuss next"

Next, we look at the atomic loss rate during the transfer process. (Fig.~\ref{f-det}?)
The atomic loss rate scales as $P^{2.58}$ but shows little to no dependency on the tweezer frequency. The lack of frequency dependency suggests that most of the atomic loss are due to the coupling to states other than $v=0$. After subtracting 0.58 from the power scaling due to the effect of confinement on the square of the atomic matrix element, the loss rate is proportional to $P^2$. This is evidence that the loss is likely caused by a two photon scattering process rather than one photon scattering. The strong power scaling causes a faster than expected loss rate for the atom. However, the atomic scattering rate contributes only a small fraction of the total scattering rate so this is not the limiting factor for the transfer.

\begin{figure*}
  \includegraphics[height=4.5cm]{fig4.pdf}
  \caption{Raman transition parameters as a function of tweezer/Raman power and detuning.
    (A) Raman Rabi frequency scales as $P_{tweezer}^{1.29}$ and follows $1/\delta$ with an offset.
    (B) Atom loss rate scales as $P_{tweezer}^{2.58}$ suggesting a two-photon scattering loss.
    (C) Molecular loss rate scales closed to $P_{tweezer}$ and follows $1/\delta^2$
    with additional detuning independent background loss rate.
    This result controdicts the our previous measurement of excited state lifetime
    suggesting that there are additional loss mechanism that is not taken into account.
    \label{f-det}}
\end{figure*}

Finally, we fit the scaling for the molecular loss rate. (Fig.~\ref{f-det}?)
We observed a dependency on the detuning that is consistent with $1/\delta^2$ suggesting that the loss is indeed mostly coming from the $v=0$ excited state (but this is only true at certain single photon detunings?). (preliminary) We observed a power scaling power of $\approx1.2$ which is closed to the prediction for a single photon scattering process. The excited state line width calculated from the scattering rate is $...(1 GHz?)$ which is much larger than the theory prediction of $10 MHz$ as well as the $20 MHz$ upper bound from our previous PA measurements (cite).

%%%% Maybe this following section about the B field can be relegated to a figure caption?
% Figure () shows the geometry of the experiment setup. A $8.8\mathrm{G}$ B field was used to defines the quantization axis in the experiment and is applied perpendicular to the tweezer axis. The tweezer beam (both frequencies for the Raman transition) has a $\pi$ polarization relative to the quantization axis, which allows us to selectively drive the Raman transition to a final state with the same total $m_F$ quantum number as our initial state.

B field dependency $... kHz/G$ which agrees with theory prediction of $... kHz/G$.
Dependency on tweezer power $... kHz/mW$, extrapulated to obtain the bare resonance at $0$ tweezer power to be at $... MHz$.

>>>>

\begin{figure*}
  \includegraphics[height=5cm]{fig1.pdf}
  \caption{Optical creation of single molecule from single atoms in tweezer.
    (A) Schematics of the optical transition from an atom pair to a weakly bound molecule.
    The initial state is the relative motional ground state between the two atoms
    and the final state is the first molecular bound state.
    The transition is driven by a pair of laser frequencies matching the binding energy
    of the molecule.
    The lasers are detuned from an excited molecular state in the $c^3\Sigma$ potential
    by $\Delta$ in order to suppress the scattering during the transfer.
    (B) Comparison between using a weakly bound and a deeply bound excited state
    as intermediate state for the Raman transition.
    The deeply bound excited state (upper half $v'=0$) has a smaller Raman Rabi frequency ($\Omega_{Raman}$)
    compared to the weakly bound excited state (lower half $v'=63$) at a given detuning.
    However, the scattering rate ($\Gamma_{scatter}$) is also much lower,
    which results in a larger Raman Rabi frequency to scattering rate ratio.
    (C) Enhancement of short range wavefunction.
    The large scattering length for the $Na(2,2),Cs(3,3)$ state creates an interaction shift
    comparable to the axial trapping frequency.
    This causes a significant change in the relative wavefunction especially at short
    intranuclear distance ($R$).
    Compared to other spin states with weaker interaction,
    the wavefunction at short distance ($R<100\mathrm{\AA}$) is significantly enhanced.
    \label{f-theory}
  }
\end{figure*}

@@

% The optical transfer scheme we use is shown in figure \todo{}. To drive the atoms from
% We use an optical Raman transition to drive the system from the atomic initial state to a ground electronic molecular state. In order to reduce the size mismatch between the atomic and molecular states, we selected the first bound state for the 3322 spin state \todo{asymtopt to the 3322 threshold?} as our final state. \todo{move selection of intermediate state to intro?}. For similar reason, the natural choice for the intermediate excited molecular state is one with highly excited motional level.However, from our calculation (and experiment?), the smaller level space for high vibrational state and the smaller detuning from the atomic threshold increases the scattering rate of the molecular state which causes a reduced Raman Rabi frequency to decoherence rate ratio and a lower transfer efficiency. Therefore, in our experiment, we selected the v'=0 state as the intermediate state for our Raman transition.

@@

To find the optimal excited intermediate state, we perform a  calculation of the Raman Rabi frequency and scattering rate at different detunings from the atomic threshold taking into account of all states of the $c^3\Sigma$ excited molecular state potential (full result in SM)~\cite{Liu2017} \textcolor{red}{KW: I have only done the calculations with c3Sigma only, cause the b3Pi is highly spin orbit coupled, and not necessarily trustworthy where we are. I can also include a citation to the paper with that potential?}. As shown in Fig.~\ref{f-theory}B, the ratio of the Raman Rabi rate to scattering rate can be made larger for more deeply bound states compared to weakly bound states at a cost of a smaller Raman Rabi frequency.  %Since for a typical molecular potential the spacing between states
% decreases with increasing vibrational levels,
% a weakly bound excited molecular state
% This further increases the scattering rate for a weakly bond excited state.
% Based on these considerations, % a weakly bound excited molecular state may not be the best choice for the intermediate state in the Raman transfer. Thus,
% we calculated both the Raman Rabi frequency and the scattering rate at different detunings (full result is shown in SM). As seen in Fig.~\ref{f-theory}B,
% we found a distinct optimal regime from what has been explored experimentally previously.
% Namely, despite a smaller Raman Rabi frequency at deeply bound intermediate states, the Rabi frequency to scattering rate ratio is much higher as compared to a weakly bound intermediate state.
As a result, we choose a deeply bound molecular excited state ($v'=0$ of $c^3\Sigma^+$) as an intermediate state to drive the Raman transition. %(mention the trend, also v'=12 will work as similarly as v'=0)

@@

The ratio of molecule scattering rate to the Rabi frequency is larger than the theory prediction
by more than a factor of $10$.
Based on the calculation above, if this comes from the coupling to the $v'=0$ excited state,
it can be either due to a high ratio of $\Omega_m / \Omega_a$ or a large $\Gamma_e$.
Additionally, coupling to other excited states can also adds an offset to both the Raman
Rabi frequency and the scattering rate which can affect the scattering rate
to Rabi frequency ratio.

In order to verify whether any one of these known sources are the origin of the discrepancy,
we measured the Raman resonance as a function of the tweezer power and single photon frequency.
The important results from the fits are the 2 photon resonance frequency (light shift),
Raman Rabi frequency, atomic lifetime and molecular lifetime.
Each of these quantities provide us with information about a different combination
of physical processes and the detuning dependency of them allows us
to distinguish the contribution from the $v'=0$ excited states and other states.

First we look at the change in resonance frequency.
As a function of the tweezer power, we observe a linear dependency on the resonance frequency
caused by the differential light shift between the atomic and molecular state (Fig.~\ref{f-det}A).
When we vary the tweezer frequency around the $v'=0$ excited state,
we can further observe a $1/\delta$ component and a constant component
in the experimentally explored region.
The background is caused by coupling to other excited states that are further away in energy.
The $1/\delta$ component, however, is predominantly due to the coupling between
the molecular state and the $v'=0$ excited state.
From this measurement, we can calculate a matrix element between the molecular state and
the excited state, $\Omega_m$, of $2\pi\times89.40(17) \mathrm{MHz}/\sqrt{\mathrm{mW}}$
or $2\pi\times346.24(65) \mathrm{MHz}$ for the $15 \mathrm{mW}$ tweezer power used above.
This number is about $3$ times larger than the value of
$2\pi\times27 \mathrm{MHz}/\sqrt{\mathrm{mW}}$ \todo{ref/sm theory}.
\todo{make sure theory part mentions $\Omega_{down}\gg\Omega_{up}$,
  also $\Omega$ vs $\Omega'$ for the single leg vs cross coupling number.}

In order to calculate the matrix element ratio,
next we discuss the dependencies of the Raman Rabi frequency.
The Raman Rabi frequency shows a non-linear dependency on the tweezer power due to the change
in the atomic wavefunction caused by the change in confinement (Fig.~\ref{f-det}B).
The atomic matrix element, $\Omega_a$,
is proportional to the the short range atomic wavefunction amplitude,
which scales as $P^{0.375}$ for weakly interacting particles.
However, due to the strong interaction between the two atoms, this approximation breaks down.
Instead, theory calculation shows that the scaling is very well approximated by $P^{0.29}$
within the range of confinement in our experiment.
Combined with the intensity factor, the Raman Rabi frequency should scale with $P^{1.29}$,
which agrees with our experimental result.
Similar to the light shift, there is also an offset and
a $v'=0$ component in the Raman Rabi frequency.
Based on the $v'=0$ component and the $\Omega_m$ measured above,
we measured a up leg Rabi frequency of $...$, or $...$ at $15 \mathrm{mW}$ tweezer power.
This gives a Rabi frequency, and therefore matrix element, ratio of $...$,
which is in fact better than the theory prediction of $...$
and should not cause the scattering from the $v'=0$ state to be higher than expected.

Unfortunately, the offset in the Raman Rabi frequency fit cancels
the Rabi frequency for red detuning Raman transition and reduces it by about $...\times$
at the current detuning. However, this difference is not enough to explain
the difference we observed in the experiment.
The same offset will increase the Rabi frequency for blue detuning
but we have observed additional excited states at slightly higher frequencies
which prevent the blue side of the transition to be usuable for Raman transition.

\begin{figure*}
  \includegraphics[height=4.5cm]{fig4.pdf}
  \caption{Raman transition parameters as a function of tweezer and Raman power and detuning.
    (A) The light shift of the Raman resonance scales as $P_{tweezer}$
    and follows $1/\delta$ with an offset.
    (B) Raman Rabi frequency scales as $P_{tweezer}^{1.29}$ and follows $1/\delta$ with an offset.
    From these results we can confirm the theory prediction of
    the atom-molecule matrix element ratio.
    (C) Atomic scattering rate scales as $P_{tweezer}^{2.58}$,
    this is consistent with a two photon scattering process.
    We have not measured a clear dependency of the loss rate on the tweezer detuning.
    \label{f-det}}
\end{figure*}

Based on our previous measurements of the excited state using photoassociation (PA) spectroscopy,
the linewidth of the $v'=0$ excited state is no larger than $50 \mathrm{MHz}$.
This suggests that the excited state linewidth should not cause a stronger than expected
scattering from $v'=0$ state either.

These results suggests that the decoherence or loss we observed during the the Raman transition
comes from either a higher than expected background scattering rate
or a different intrinsic or technical source that we have not accounted for.
As mentioned above, we have observed significant improvment in the coherence time
by filtering out most of the ASE from the laser suggesting the spectral purity of
the laser is a significant source of scattering.
Although we have good control on the total power of the tweezer and Raman beam to stabilize
the light shift on the transition, the power ratio of the two frequencies within the beam
is not yet controlled as tightly and may result in a residual resonance fluctuation.
Finally, Fig.~\ref{f-det}C shows the two-body scattering rate for the atomic initial state.
The scattering rate scales as $P_{tweezer}^{2.58}$ which is inconsistent with
the prediction based on single photon scattering process.
Although the absolute scattering rate is much lower than the total decoherence rate
and we have not been able to observe a dependency on the detuning in order to
verify if the scattering process is related to the $v'=0$ state, the power scaling
strongly suggests the existence of a unknown two photon scattering process.

%%%% Maybe this following section about the B field can be relegated to a figure caption?
% Figure () shows the geometry of the experiment setup. A $8.8\mathrm{G}$ B field was used to defines the quantization axis in the experiment and is applied perpendicular to the tweezer axis. The tweezer beam (both frequencies for the Raman transition) has a $\pi$ polarization relative to the quantization axis, which allows us to selectively drive the Raman transition to a final state with the same total $m_F$ quantum number as our initial state.
B field dependency $... kHz/G$ which agrees with theory prediction of $... kHz/G$.
Dependency on tweezer power $... kHz/mW$, extrapulated to obtain the bare resonance at $0$ tweezer power to be at $... MHz$.
